Electric polarizability as the fundamental variable in the dielectric properties of two-dimensional materials

### 2019-01-04 SI.tex
Add changes to SI

\subsection{Profile of induced dipoles of 2D material}
\label{sec:dipole-plot}
Here we show in detail the
$\Delta {\rho}=\rho(\boldsymbol{E}) - \rho(\boldsymbol{E}=0)$ profile of
the 2H-MoS$_{2}$ slab in main text Figure \ref{main-fig-1}. The
density $\Delta \rho$ is calculated via $\Delta \rho(z) = \frac{1}{S} \int_{S} \Delta \rho (x,y;z) dx dy $, 
where $S$ is the surface of the unit cell perpendicular to the given direction, in this case $z$. 
%
As can be seen in Figure \ref{fig:rho-profile} the induced charges on the
MoS$_{2}$ layer only extends to a width of $\sim{}$12 \AA{} centered at the middle of the layer. This corresponds to about 6 \AA ~from each side. 
When the
SL size $L \gg$12 \AA{} as in the first principle calculations done in
the main text, the induced dipoles from the periodic images do not
interact, and thus give the converged values of $\alpha_{\mathrm{2D}}$.

@@

We calculate the frequency-dependent dielectric properties of BN with
varying supperlattice $L$ at PBE, G$_{0}$W$_{0}$, and G$_{0}$W$_{0}$--BSE. 
Figure \ref{fig:PBE-omega-in} and \ref{fig:PBE-omega-out} show the
magnitudes of $\varepsilon_{\mathrm{SL}}^{\parallel}(\omega)$ and
$\varepsilon_{\mathrm{SL}}^{\perp}(\omega)$ reduce with increasing $L$
throughout the whole frequency range at PBE level. The $L$-dependency can also be
removed using Eqs. \ref{main-eq:alpha-para-def} and
\ref{main-eq:alpha-perp-def}, yielding lattice-independent
polarizabilities throughout the frequency domain (Figure \ref{fig:PBE-omega-in}{\bf b}-{\bf d} and \ref{fig:PBE-omega-out}{\bf b}-{\bf d}). 
%
Note that when extracting
the frequency-dependent 2D polarizabilities from
Eqs. \ref{main-eq:alpha-para-def} and \ref{main-eq:alpha-perp-def},
the peak position in energy $\omega (eV)$ from $\varepsilon_{\mathrm{SL}}$ is preserved. This
is explained by the fact that the local extrema from spectra of
$\varepsilon_{\mathrm{SL}}$ is also the local extrema in corresponding
$\alpha_{\mathrm{2D}}$, since when
$\partial \varepsilon_{\mathrm{SL}}(\omega) / \partial \omega = 0 $,
we have
\begin{subequations}
\begin{eqnarray}
  \label{eq:extrema-para}
  \frac{\partial \alpha_{\mathrm{2D}}^{\parallel}(\omega)}{\partial \omega}
  &= \varepsilon_{0}L {\displaystyle \frac{\partial \varepsilon_{\mathrm{SL}}^{\parallel}}{\partial \omega}} = 0   \\
  \label{eq:extrema-perp}
  \frac{\partial \alpha_{\mathrm{2D}}^{\perp}(\omega)}{\partial \omega}
  &= \varepsilon_{0}L {\displaystyle \frac{1}{\varepsilon_{\mathrm{SL}}^{2}}\frac{\partial \varepsilon_{\mathrm{SL}}^{\parallel}}{\partial \omega}} = 0
\end{eqnarray}
\end{subequations}
which indicates that no corrections in energy are present 
when transforming
$\varepsilon_{\mathrm{SL}}(\omega)$ to $\alpha_{\mathrm{2D}}(\omega)$.

We note that the result change slightly when G$_{0}$W$_{0}$ method
is used. As shown in Figures \ref{fig:GW-omega-in} and
\ref{fig:GW-omega-out}, not only do the intensity of the dielectric
functions change with $L$, but also the peak positions. As a result
the obtained polarizabilities also show $L$-dependent peak shift. This
can be explained by the long-range nature of Coulomb interactions. In
fact the self-energy of the G$_{0}$W$_{0}$ system $\Sigma$, is determined by:
\begin{equation}
  \label{eq:GW-formula}
  \Sigma = i G_{0} W_{0} = i G_{0} \varepsilon^{-1} \nu
\end{equation}
where $G$ is the non-interacting Green function, and $\nu$ is the
Coulomb interaction. Since the dielectric constant $\varepsilon$
changes with $L$, the self energy as well quasi-particle energies of
the G$_{0}$W$_{0}$ method also varies with $L$. As a result we get
higher peak energy of the dielectric spectrum when $L$ increases. Such
peculiar $L$-dependency can be minimized when adding electron-hole
interaction through the Bethe-Salpethe equation (BSE), as shown in
Figures \ref{fig:BSE-omega-in} and \ref{fig:BSE-omega-out}. The
polarizabilities again becomes lattice-independent due to the
localization of excitons within the 2D material layer, which corrects the energy shift of G$_{0}$W$_{0}$ method.

@@

\subsection{Detailed derivations of Eq. \ref{main-eq:2D-Moss-para}}  %
\label{ssec:theory-1-para}

In this section we show how Eq. \ref{main-eq:alpha-para-def}
is derived from Eq. \ref{eq:alpha-RPA}. For the in-plane component
$\alpha_{\mathrm{2D}}^{\parallel}$, $e^{-i\mathbf{qr}}$ is independent
of $z$, therefore the integral in
$|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}$
becomes independent of $\psi^{\perp}$, due to the orthogonality
and normality. The Bloch-wave form of $\psi^{\parallel}$ ensures that
only the cell-function $u(\mathbf{k})$ contributes to the final result
of $\alpha^{\parallel}_{\mathrm{2D}}$ \cite{davies_physics_1997}, such
that:
\begin{equation}
  \label{eq:alpha_para_RPA}
  \alpha_{\mathrm{2D}}^{\parallel} = \frac{2e^{2}}
  {(2 \pi)^{2}} \int d^{2}\mathbf{k} \sum_{\mathrm{c, v}}
  \frac{|<u_{\mathrm{c}}(\mathbf{k})|\nabla|u_{\mathrm{v}}(\mathbf{k})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
Following the method of $\mathbf{k} \cdot \mathbf{p}$ theory from
Ref. \citenum{Jiang_2017_Eg_Eb}, the matrix element in the numerator
of Eq. \ref{eq:alpha_para_RPA} is approximated by:

@@

The coefficient of $C^{\parallel}$ adapted from
Ref. \citenum{Jiang_2017_Eg_Eb} $C^{\parallel} = N e^{2}/(2\pi)$
predicts linear correlation between $\alpha^{\parallel}_{\mathrm{2D}}$
and $E_{\mathrm{g}}^{-1}$. We validate this by examining the
DFT-calculated
$(4\pi \varepsilon_{0})/\alpha^{\parallel}_{\mathrm{2D}}$ (measured in
\AA{}$^{-1}$) and $E_{\mathrm{g}}$ (measured in eV) in
Figure \ref{main-fig-3}\textbf{c}. The coefficient $C^{\parallel}$
becomes
$8 \pi^{2} \varepsilon_{0} \AA{} / (e N) \approx 0.436 / N = 0.183$,
corresponding to $N$ between 2 and 3, which is a reasonable result for
the 2D materials studied.

@@

\subsection{Detailed derivation of Eq. \ref{main-eq:2D-Moss-perp}}
\label{ssec:theory-1-perp}

For Eq. \ref{main-eq:2D-Moss-perp}, treating the in-plane
wave functions as plane wave with form
$\psi^{\parallel}(\rho) \propto e^{i \mathbf{k \rho}}$, the matrix
element of
$<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>$,
when $\mathbf{q}=(0, 0, q_{z})$, becomes\cite{Hybertsen_1987}:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z}
  <\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>
  &= \frac{1}{A} \int dx \int dy
  e^{i(\mathbf{-k \rho} - \mathbf{q \rho} + \mathbf{(k+q) \rho})}
  \int (\psi^{\perp})^{*}_{\mathrm{v}}(\mathbf{k})e^{-iq_{z}z}\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})\\
  &= <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})>
  \end{aligned}
\end{equation}
Note that the states perpendicular are bound, the integral is
meaningful only when $\mathbf{k=k+q}$ \cite{davies_physics_1997}. By
performing the Taylor expansion of
$e^{-i\mathbf{qr}} \approx 1 - i\mathbf{qr}$, we get:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z-2}
    <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
    &\approx <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})> -
    iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>\\
    &= -iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
   \end{aligned}
\end{equation}
plug this into Eq. \ref{eq:alpha-RPA} and express the summation over
$k_{x}$ and $k_{y}$ in a continuous form within the Brillouin Zone, we
arrive at:
\begin{equation}
\label{eq:alpha_perp_RPA}
\alpha_{\mathrm{2D}}^{\perp} = \frac{2e^{2}}{(2 \pi) ^{2}} \int d^{2}\mathbf{k}
\sum_{\mathrm{c, v}}
\frac{|<\psi_{\mathrm{v}}(\mathbf{k})|z|\psi_{\mathrm{c}}(\mathbf{k})>|^{2}}
{E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
The formalism is slightly different from Eq.\ref{eq:alpha_para_RPA}.

@@

\subsection{Explanation of Eq. \ref{main-eq:2D-Moss-perp}
  from fundamental electrostatics}
\label{ssec:theory-1-perp-fundamental}

The dependency of $\alpha_{\mathrm{2D}}^{\perp}$ on the thickness
$\delta$ of a 2D material, can also be regarded using fundamental
electrostatic model. Consider the smallest repeating unit of the 2D
material with xy-plane area $A$, under small perturbation field $E$
along the z-direction.  Note that the surface bound charge
$\sigma_{\mathrm{b}}=n e /A$, where $n$ is the number of unit charges
contributes to the bound charges, comes only from the dipoles of the
outer-most atoms, since the induced charges from inner atoms are
cancel out (see Figure \ref{fig:classic-model}).
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.99\linewidth]{img/SI-classic.pdf}
  \caption{Fundamental electrostatic model for the
    thickness-dependency of $\alpha_{\mathrm{2D}}^{\perp}$, using 2H-MoS2 as an
    example. Left: induced dipoles from individual atoms along the
    z-direction. The positive and negative induced charges from inner
    atoms cancel out. Right: simplified model for the thickness
    dependency of $\alpha_{\mathrm{2D}}^{\perp}$, where the surface dipole density
    $\boldsymbol{\mu}$ comes only from the outer-most atoms.}
  \label{fig:classic-model}
\end{figure}
From the definition of
$\alpha_{\mathrm{2D}}^{\perp}$, we have:
\begin{equation}
  \label{eq:alpha-classic}
  \alpha_{\mathrm{2D}}^{\perp} = \frac{\boldsymbol{u}_{z}}{E_{\mathrm{loc}} A}
  = \frac{(d_{\mathrm{max}} + r_{\mathrm{cov}}^{i} + r_{\mathrm{cov}}^{j}) \sigma_{\mathrm{b}}}{E}
\end{equation}
where $r_{\mathrm{cov}}^{i}$ and $r_{\mathrm{cov}}^{j}$ are the
covalent radii of the outer-most atoms, the characteristic length of
the dipole extension in z-direction, respectively, and $d_{\mathrm{max}}$ is the
z-distance between the nuclei of such atoms.  The field $E$
counterbalances the field from the surface bound charges and equals
$E = \sigma_{\mathrm{b}}/\varepsilon_{0}$. Therefore we have:
\begin{equation}
  \label{eq:alpha-classic-2}
  \alpha_{\mathrm{2D}}^{\perp} = (d_{\mathrm{max}} + r_{\mathrm{cov}}^{i} + r_{\mathrm{cov}}^{j})\varepsilon_{0}
                = \delta_{\mathrm{cov}} \varepsilon_{0}
\end{equation}
which explains the linear relation seen in Figure
\ref{main-fig-3}d. We can see that such simple model nicely captures
the thickness feature of $\alpha_{\mathrm{2D}}^{\perp}$, and
reproduces the right coefficient between $\delta_{\mathrm{cov}}$ and
$\alpha_{\mathrm{2D}}^{\perp}$.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1.05\linewidth]{img/SI-gpaw-data.pdf}
  \caption{Validation of the linear relation between \textbf{a,}
    $1/\alpha_{\mathrm{2D}}^{\parallel}-E_{\mathrm{g}}(\mathrm{HSE})$
    and \textbf{b,}
    $\alpha_{\mathrm{2D}}^{\perp}-\delta_{\mathrm{cov}}$ from 
    Ref.\cite{Haastrup_2018} corresponding to Figure \ref{main-fig-3}{\bf c}
    and \ref{main-fig-3}{\bf d}.}
  \label{fig:gpaw-alpha-relation}
\end{figure}

Due to time-consuming simulations and significant increment in 
memory overload, high-accurate calculations at hybrid HSE06 level is limited 
to about 56 compounds. 
It is desirable to validate our proposed
relations on an even larger scale database. We select over 230
semiconducting 2D materials from Ref.\cite{Haastrup_2018} 
with a GW bandgap
larger than 0.05 eV and extracted the 2D polarizabilities calculated
on the PBE level. The proposed linear relations between
$1/\alpha_{\mathrm{2D}}^{\parallel}-E_{\mathrm{g}}(\mathrm{HSE})$ and
$\alpha_{\mathrm{2D}}^{\perp}-\delta_{\mathrm{cov}}$ are also valid,
as shown in Figure \ref{fig:gpaw-alpha-relation}. Excellent linear
correlation is observed in both cases with the $R^{2}$ coefficient
larger than 0.9 which indicates the existence of a universal description 
of 2D dielectric nature through the proposed relations with the 2D
polarizabilities. We note that the slope of the linear regression is
slightly different from the one proposed from the dielectric
response at the HSE06 level.

@@

\subsection{Explanation for the separation between 2D and 3D regimes in Figure \ref{main-fig:aniso}}

>>>>

\subsection{Profile of induced dipoles of 2D material}
\label{sec:dipole-plot}
Here we show in detail the
$\Delta {\rho}=\rho(\boldsymbol{E}) - \rho(\boldsymbol{E}=0)$ profile of
the 2H-MoS$_{2}$ slab in main text Figure \ref{main-fig-1}. The
density $\Delta \rho$ is calculated via $\Delta \rho(z) = \frac{1}{S} \int_{S} \Delta \rho (x,y;z) dx dy $, 
where $S$ is the surface of the unit cell perpendicular to the given direction, in this case $z$. 
%
As can be seen in Figure \ref{fig:rho-profile} the induced charges on
the MoS$_{2}$ layer only extends to a width of $\sim{}$12 \AA{}
centered at the middle of the layer. This corresponds to about 5-6 \AA
from each side.  When the SL size $L \gg$12 \AA{} as in the first
principle calculations done in the main text, the induced dipoles from
the periodic images do not interact, and thus give the converged
values of $\alpha_{\mathrm{2D}}$.

@@

We calculate the frequency-dependent dielectric properties of BN with
varying supperlattice $L$ at PBE, G$_{0}$W$_{0}$, and G$_{0}$W$_{0}$--BSE. 
Figure \ref{fig:PBE-omega-in} and \ref{fig:PBE-omega-out} show the
magnitudes of $\varepsilon_{\mathrm{SL}}^{\parallel}(\omega)$ and
$\varepsilon_{\mathrm{SL}}^{\perp}(\omega)$ reduce with increasing $L$
throughout the whole frequency range at PBE level. The $L$-dependency can also be
removed using main text Eqs. \ref{main-eq:alpha-para-def} and
\ref{main-eq:alpha-perp-def}, yielding lattice-independent
polarizabilities throughout the frequency domain (Figure \ref{fig:PBE-omega-in}{\bf b}-{\bf d} and \ref{fig:PBE-omega-out}{\bf b}-{\bf d}). 
%
Note that when extracting the frequency-dependent 2D polarizabilities
from main text Eqs. \ref{main-eq:alpha-para-def} and
\ref{main-eq:alpha-perp-def}, the peak position in energy
$\hbar \omega$ from $\varepsilon_{\mathrm{SL}}$ is preserved. This is
explained by the fact that the local extrema from spectra of
$\varepsilon_{\mathrm{SL}}$ is also the local extrema in corresponding
$\alpha_{\mathrm{2D}}$, since when
$\partial \varepsilon_{\mathrm{SL}}(\omega) / \partial \omega = 0 $,
we have
\begin{subequations}
\begin{eqnarray}
  \label{eq:extrema-para}
  \frac{\partial \alpha_{\mathrm{2D}}^{\parallel}(\omega)}{\partial \omega}
  &= \varepsilon_{0}L {\displaystyle \frac{\partial \varepsilon_{\mathrm{SL}}^{\parallel}}{\partial \omega}} = 0   \\
  \label{eq:extrema-perp}
  \frac{\partial \alpha_{\mathrm{2D}}^{\perp}(\omega)}{\partial \omega}
  &= \varepsilon_{0}L {\displaystyle \frac{1}{\varepsilon_{\mathrm{SL}}^{2}}\frac{\partial \varepsilon_{\mathrm{SL}}^{\parallel}}{\partial \omega}} = 0
\end{eqnarray}
\end{subequations}
which indicates that no corrections in energy are present 
when transforming
$\varepsilon_{\mathrm{SL}}(\omega)$ to $\alpha_{\mathrm{2D}}(\omega)$.
Performing the simulations at the level of G$_{0}$W$_{0}$
we observe shifts in energy in $\varepsilon_{\mathrm{SL}}(\omega)$ 
as a function of $L$. Figures \ref{fig:GW-omega-in} and
\ref{fig:GW-omega-out} show that not only the magnitude of the dielectric
functions change with $L$ but also the peak positions. As a result
the obtained polarizabilities also show $L$-dependent peak 
shift (Figures \ref{fig:GW-omega-in} {\bf b}-{\bf d} and
\ref{fig:GW-omega-out}{\bf b}-{\bf d}). This
can be explained by the long-range nature of Coulomb interactions into the self-energy 
$\Sigma=iGW$.

%%%%%%%%%%%%%%%%%%%%%%%
The non-interacting Green's function can be constructed as:

\begin{equation}
    G^{(0)}(\mathbf{r},\mathbf{r}',\omega)=\sum_n\frac{\phi_{\mathrm{n}\mathbf{k}}^{(0)}(\mathbf{r})\phi_{\mathrm{n}\mathbf{k}}^{*(0)}(\mathbf{r'})}{\omega-\epsilon_\mathrm{n}-+i\eta\mathrm{sgn}(\epsilon_\mathrm{n}-\epsilon_\mathrm{F})}
\end{equation}

Here $\epsilon_n(\mathbf{k})$ are the DFT eigenenergies at \textbf{k}, $\epsilon_F$ the Fermi energy, $\omega$ the frequency, $\ket{u_{n\mathbf{k}}}$ the cell periodic Bloch functions, $\phi_{n\mathbf{k}}$ are the one-electron orbitals and $\eta$ is an infintesimal complex shift. It can be seen from equation \ref{eq:dft-dielectric} there is a clear volume dependence on $\Omega$ of the dielectric function from the preceding DFT calculation. Thus, when carrying out a calculation on a slab, the dielectric function will vary with the vacuum spacing used.

Using the dielectric function we can calculate the screened Coulomb interaction:
\begin{equation}
    W=\epsilon^{-1}\nu
\end{equation}
where $\nu$ is the bare Coulomb interaction given by
$e^2$/$|\mathbf{r}-\mathbf{r}'|$. Due to the
1/$|\mathbf{r}-\mathbf{r}'|$ term, images in the non-periodic
direction have a long-range spurious interaction which varies with
vacuum spacing. From the screened Coulomb interaction and
non-interacting Green's function it is possible to calculate the
self-energy of the system:
\begin{equation}\label{self}
\Sigma=iGW
\end{equation}
and the quasi-particle eigenenergies are found using:
\begin{equation}\label{quasi}
E^{\mathrm{QP}}_{\mathrm{n}\mathbf{k}}=\Re\Big[\braket{\phi_{\mathrm{n}\mathbf{k}}|-\frac{1}{2}\Delta+V_{\mathrm{ext}}+V_\mathrm{H}+\Sigma(\epsilon_{\mathrm{n}\mathbf{k}}^{\mathrm{DFT}})|\phi_{\mathrm{n}\mathbf{k}}}\Big]
\end{equation}
Therefore a new set of eigenergies, E$_{n\mathbf{k}}^{QP}$ are found. Using E$_{n\mathbf{k}}^{QP}$ it is then possible to recalculate the dielectric function using the quasi-particle eigenergies as well as the DFT eigenfunctions.

Due to the volume dependence of $\Sigma$ in equations \ref{self} and
\ref{quasi}, increasing the vacuum spacing leads to a change in the
calculated quasi-particle eigenergies. This, along with the volume
dependence in the calculation of the dielectric function, leads to an
increase in energy of the peak position and a decrease in its
magnitude with increasing vacuum spacing.  Including excitonic effects
on top of a G$_0$W$_0$ calculation through solving the Bethe--Salpeter
equation corrects for this energy shift by localizing the exciton
within the slab (Figures \ref{fig:BSE-omega-in} and
\ref{fig:BSE-omega-out}). However, the decrease in the magnitude of
the dielectric function on both components of
$\varepsilon_{\mathrm{SL}}(\omega)$ with increasing vacuum spacing is
still observed.  Using main text Eqs. \ref{main-eq:alpha-para-def} and
\ref{main-eq:alpha-perp-def} we can remove the dependence on $L$ as
plotted in Figures \ref{fig:BSE-omega-in}{\bf b}-{\bf d} and
\ref{fig:BSE-omega-out}{\bf b}-{\bf d} with few variations on
$\varepsilon^{\perp}_{\mathrm{SL}}(\omega)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% We note that the result change slightly when G$_{0}$W$_{0}$ method                                               %
% is used. As shown in Figures \ref{fig:GW-omega-in} and                                                           %
% \ref{fig:GW-omega-out}, not only do the intensity of the dielectric                                              %
% functions change with $L$, but also the peak positions. As a result                                              %
% the obtained polarizabilities also show $L$-dependent peak shift. This                                           %
% can be explained by the long-range nature of Coulomb interactions. In                                            %
% fact the self-energy of the G$_{0}$W$_{0}$ system $\Sigma$, is determined by:                                    %
% \begin{equation}                                                                                                 %
%   \label{eq:GW-formula}                                                                                          %
%   \Sigma = i G_{0} W_{0} = i G_{0} \varepsilon^{-1} \nu                                                          %
% \end{equation}                                                                                                   %
% where $G$ is the non-interacting Green function, and $\nu$ is the                                                %
% Coulomb interaction. Since the dielectric constant $\varepsilon$                                                 %
% changes with $L$, the self energy as well quasi-particle energies of                                             %
% the G$_{0}$W$_{0}$ method also varies with $L$. As a result we get                                               %
% higher peak energy of the dielectric spectrum when $L$ increases. Such                                           %
% peculiar $L$-dependency can be minimized when adding electron-hole                                               %
% interaction through the Bethe-Salpethe equation (BSE), as shown in                                               %
% Figures \ref{fig:BSE-omega-in} and \ref{fig:BSE-omega-out}. The                                                  %
% polarizabilities again becomes lattice-independent due to the                                                    %
% localization of excitons within the 2D material layer, which corrects the energy shift of G$_{0}$W$_{0}$ method. %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Detailed derivations of main text Eq. \ref{main-eq:2D-Moss-para}}  %
\label{ssec:theory-1-para}

In this section we show how Eq. \ref{main-eq:alpha-para-def}
is derived from Eq. \ref{eq:alpha-RPA}. For the in-plane component
$\alpha_{\mathrm{2D}}^{\parallel}$, $e^{-i\mathbf{qr}}$ is independent
of $z$, therefore the integral in
$|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}$
becomes independent of $\psi^{\perp}$, due to the orthogonality
and normality. The Bloch-wave form of $\psi^{\parallel}$ ensures that
only the cell-function $u(\mathbf{k})$ contributes to the final result
of $\alpha^{\parallel}_{\mathrm{2D}}$ \cite{davies_physics_1997}, such
that:
\begin{equation}
  \label{eq:alpha_para_RPA}
  \alpha_{\mathrm{2D}}^{\parallel} = \frac{2e^{2}}
  {(2 \pi)^{2}} \int d^{2}\mathbf{k} \sum_{\mathrm{c, v}}
  \frac{|<u_{\mathrm{c}}(\mathbf{k})|\nabla|u_{\mathrm{v}}(\mathbf{k})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
Following the method of $\mathbf{k} \cdot \mathbf{p}$ theory from
Ref.~\citenum{Jiang_2017_Eg_Eb}, the matrix element in the numerator
of Eq. \ref{eq:alpha_para_RPA} is approximated by:

@@

The coefficient of $C^{\parallel}$ adapted from
Ref.~\citenum{Jiang_2017_Eg_Eb} $C^{\parallel} = N e^{2}/(2\pi)$
predicts linear correlation between $\alpha^{\parallel}_{\mathrm{2D}}$
and $E_{\mathrm{g}}^{-1}$. We validate this by examining the
DFT-calculated
$(4\pi \varepsilon_{0})/\alpha^{\parallel}_{\mathrm{2D}}$ (measured in
\AA{}$^{-1}$) and $E_{\mathrm{g}}$ (measured in eV) in
Figure \ref{main-fig-3}\textbf{c}. The coefficient $C^{\parallel}$
becomes
$8 \pi^{2} \varepsilon_{0} \AA{} / (e N) \approx 0.436 / N = 0.183$,
corresponding to $N$ between 2 and 3, which is a reasonable result for
the 2D materials studied.

@@

\subsection{Detailed derivation of main text Eq. \ref{main-eq:2D-Moss-perp}}
\label{ssec:theory-1-perp}

For main text Eq. \ref{main-eq:2D-Moss-perp}, treating the in-plane
wave functions as plane wave with form
$\psi^{\parallel}(\rho) \propto e^{i \mathbf{k \rho}}$, the matrix
element of
$<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>$,
when $\mathbf{q}=(0, 0, q_{z})$, becomes\cite{Hybertsen_1987}:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z}
  <\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>
  &= \frac{1}{A} \int dx \int dy
  e^{i(\mathbf{-k \rho} - \mathbf{q \rho} + \mathbf{(k+q) \rho})}
  \int (\psi^{\perp})^{*}_{\mathrm{v}}(\mathbf{k})e^{-iq_{z}z}\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})\\
  &= <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})>
  \end{aligned}
\end{equation}
Note that the states perpendicular are bound, the integral is
meaningful only when $\mathbf{k=k+q}$ \cite{davies_physics_1997}. By
performing the Taylor expansion of
$e^{-i\mathbf{qr}} \approx 1 - i\mathbf{qr}$, we get:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z-2}
    <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
    &\approx <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})> -
    iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>\\
    &= -iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
   \end{aligned}
\end{equation}
plug this into Eq. \ref{eq:alpha-RPA} and express the summation over
$k_{x}$ and $k_{y}$ in a continuous form within the Brillouin Zone, we
arrive at:
\begin{equation}
\label{eq:alpha_perp_RPA}
\alpha_{\mathrm{2D}}^{\perp} = \frac{2e^{2}}{(2 \pi) ^{2}} \int d^{2}\mathbf{k}
\sum_{\mathrm{c, v}}
\frac{|<\psi_{\mathrm{v}}(\mathbf{k})|z|\psi_{\mathrm{c}}(\mathbf{k})>|^{2}}
{E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
The formalism is slightly different from Eq.\ref{eq:alpha_para_RPA}.

@@

\subsection{Explanation of main text Eq. \ref{main-eq:2D-Moss-perp}
  from fundamental electrostatics}
\label{ssec:theory-1-perp-fundamental}

The dependency of $\alpha_{\mathrm{2D}}^{\perp}$ on the thickness
$\delta$ of a 2D material, can also be regarded using fundamental
electrostatic model. Consider the smallest repeating unit of the 2D
material with xy-plane area $A$, under small perturbation field $E$
along the z-direction.  Note that the surface bound charge
$\sigma_{\mathrm{b}}=n e /A$, where $n$ is the number of unit charges
contributes to the bound charges, comes only from the dipoles of the
outer-most atoms, since the induced charges from inner atoms are
cancel out (see Figure \ref{fig:classic-model}).
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.99\linewidth]{img/SI-classic.pdf}
  \caption{Fundamental electrostatic model for the
    thickness-dependency of $\alpha_{\mathrm{2D}}^{\perp}$, using 2H-MoS2 as an
    example. Left: induced dipoles from individual atoms along the
    z-direction. The positive and negative induced charges from inner
    atoms cancel out. Right: simplified model for the thickness
    dependency of $\alpha_{\mathrm{2D}}^{\perp}$, where the surface dipole density
    $\boldsymbol{\mu}$ comes only from the outer-most atoms.}
  \label{fig:classic-model}
\end{figure}
From the definition of
$\alpha_{\mathrm{2D}}^{\perp}$, we have:
\begin{equation}
  \label{eq:alpha-classic}
  \alpha_{\mathrm{2D}}^{\perp} = \frac{\boldsymbol{u}_{z}}{E_{\mathrm{loc}} A}
  = \frac{(d_{\mathrm{max}} + r_{\mathrm{cov}}^{i} + r_{\mathrm{cov}}^{j}) \sigma_{\mathrm{b}}}{E}
\end{equation}
where $r_{\mathrm{cov}}^{i}$ and $r_{\mathrm{cov}}^{j}$ are the
covalent radii of the outer-most atoms, the characteristic length of
the dipole extension in z-direction, respectively, and $d_{\mathrm{max}}$ is the
z-distance between the nuclei of such atoms.  The field $E$
counterbalances the field from the surface bound charges and equals
$E = \sigma_{\mathrm{b}}/\varepsilon_{0}$. Therefore we have:
\begin{equation}
  \label{eq:alpha-classic-2}
  \alpha_{\mathrm{2D}}^{\perp} = (d_{\mathrm{max}} + r_{\mathrm{cov}}^{i} + r_{\mathrm{cov}}^{j})\varepsilon_{0}
                = \delta_{\mathrm{cov}} \varepsilon_{0}
\end{equation}
which explains the linear relation seen in main text Figure
\ref{main-fig-3}d. We can see that such simple model nicely captures
the thickness feature of $\alpha_{\mathrm{2D}}^{\perp}$, and
reproduces the right coefficient between $\delta_{\mathrm{cov}}$ and
$\alpha_{\mathrm{2D}}^{\perp}$.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1.05\linewidth]{img/SI-gpaw-data.pdf}
  \caption{Validation of the linear relation between \textbf{a,}
    $1/\alpha_{\mathrm{2D}}^{\parallel}-E_{\mathrm{g}}(\mathrm{HSE})$
    and \textbf{b,}
    $\alpha_{\mathrm{2D}}^{\perp}-\delta_{\mathrm{cov}}$ from 
    Ref.~\citenum{Haastrup_2018} corresponding to main text Figure \ref{main-fig-3}{\bf c}
    and \ref{main-fig-3}{\bf d}.}
  \label{fig:gpaw-alpha-relation}
\end{figure}

Due to time-consuming simulations and significant increment in 
memory overload, high-accurate calculations at hybrid HSE06 level is limited 
to about 56 compounds. 
It is desirable to validate our proposed
relations on an even larger scale database. We select over 230
semiconducting 2D materials from Ref.~\citenum{Haastrup_2018} 
with a GW bandgap
larger than 0.05 eV and extracted the 2D polarizabilities calculated
on the PBE level. The proposed linear relations between
$1/\alpha_{\mathrm{2D}}^{\parallel}-E_{\mathrm{g}}(\mathrm{HSE})$ and
$\alpha_{\mathrm{2D}}^{\perp}-\delta_{\mathrm{cov}}$ are also valid,
as shown in Figure \ref{fig:gpaw-alpha-relation}. Excellent linear
correlation is observed in both cases with the $R^{2}$ coefficient
larger than 0.9 which indicates the existence of a universal description 
of 2D dielectric nature through the proposed relations with the 2D
polarizabilities. We note that the slope of the linear regression is
slightly different from the one proposed from the dielectric
response at the HSE06 level.

@@

\subsection{Explanation for the separation between 2D and 3D regimes in main text Figure \ref{main-fig:aniso}}
